Reintroduction of resistant frogs facilitates landscape-scale recovery in the presence of a lethal fungal disease

### 2023-12-23 doc/manuscript/translocation.tex
Compile versions for journal and preprint server

\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.

\templatetype{pnasmathematics} % Choose template 
% {pnasmathematics} = Template for a two-column research article
% {pnasmathematics} %= Template for a one-column mathematics article
% {pnasinvited} %= Template for a PNAS invited submission

\title{Reintroduction of resistant frogs facilitates landscape-scale recovery in the presence of a lethal fungal disease}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,b]{Roland A. Knapp}
\author[c,1]{Mark Q. Wilber} 
\author[d,e,1]{Allison Q. Byrne}
\author[f,g]{Maxwell B. Joseph}
\author[a,b]{Thomas C. Smith}
\author[d,e,h]{Andrew P. Rothstein}
\author[i]{Robert L. Grasso}
\author[d,e]{Erica Bree Rosenblum}

\affil[a]{Sierra Nevada Aquatic Research Laboratory, University of California, Mammoth Lakes, CA, 93546}
\affil[b]{Earth Research Institute, University of California, Santa Barbara, CA, 93106-3060}
\affil[c]{School of Natural Resources, University of Tennessee Institute of Agriculture, Knoxville, TN, 37996}
\affil[d]{Department of Environmental Science, Policy, and Management, University of California - Berkeley, Berkeley, CA, 94720-3114}
\affil[e]{Museum of Vertebrate Zoology, University of California - Berkeley, Berkeley, CA, 94720-3160}
\affil[f]{Earth Lab, University of Colorado, Boulder, CO, 80303}
\affil[g]{Current affiliation: Planet, San Francisco, CA, 94107}
\affil[h]{Current affiliation: Ginkgo Bioworks, Boston, MA, 02210}
\affil[i]{Resources Management and Science, Yosemite National Park, El Portal, CA, 95318}

% Please give the surname of the lead author for the running footer
\leadauthor{Knapp}

% Please add a significance statement to explain the relevance of your work
\significancestatement{Understanding how species persist despite accelerating global change is
critical for the conservation of biodiversity. Emerging infectious
diseases can have particularly devastating impacts, and few options
exist to reverse these effects. We used large-scale reintroductions of
disease-resistant individuals in an effort to recover a once-common frog
species driven to near-extinction by a disease that has decimated
amphibian biodiversity. Introduction of resistant frogs allowed
reestablishment of viable populations in the presence of disease. In
addition, resistance may be at least partially the result of natural
selection at specific immune function genes, which show evidence for
selection in recovering populations. The evolution of resistance and
reintroductions of resistant individuals could play an important role in
biodiversity conservation in our rapidly changing world.}

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{}
\authordeclaration{}
\equalauthors{\textsuperscript{1}M.W. contributed equally to this work with A.B.}
\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: roland.knapp@ucsb.edu}

% At least three keywords are required at submission. Please provide three to five keywords, separated by the pipe symbol.
\keywords{}

\begin{abstract}
Vast alteration of the biosphere by humans is causing a sixth mass
extinction, driven in part by an increase in emerging infectious
diseases. The emergence of the lethal fungal pathogen
(\emph{Batrachochytrium dendrobatidis}; ``Bd'') has devastated global
amphibian biodiversity, with hundreds of species experiencing declines
or extinctions. Given that there are no broadly applicable methods to
reverse these impacts in the wild, the future of many amphibians appears
grim. The once-common mountain yellow-legged (MYL) frog is emblematic of
amphibians threatened by Bd. Although most MYL frog populations are
extirpated following disease outbreaks, some persist and eventually
recover. Frogs in these recovering populations have increased resistance
against Bd infection, consistent with evolution of resistant genotypes
and/or acquired immunity. We conducted a 15-year landscape-scale
reintroduction study and show that frogs collected from recovering
populations and reintroduced to vacant habitats can reestablish
populations despite the presence of Bd. In addition, results from
viability modeling suggest that many reintroduced populations have a low
probability of extinction over 50 years. To better understand the role
of evolution in frog resistance, we compared the genomes of MYL frogs
from Bd-naive and recovering populations. We found substantial
differences between these categories, including changes in immune
function loci that may confer increased resistance, consistent with
evolutionary changes in response to Bd exposure. These results provide a
rare example of how reintroduction of resistant individuals can allow
the landscape-scale recovery of disease-impacted species. This example
has broad implications for the many taxa worldwide that are threatened
with extinction by novel pathogens.
\end{abstract}

\dates{This manuscript was compiled on \today}
% \doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\begin{document}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

\dropcap{H}uman activities are increasingly impacting global biodiversity
\citep{ceballos2015}, with important implications for ecosystem
resilience and human welfare \citep{naeem2009}. One consequence of human
alteration of the biosphere is an increase in emerging infectious
diseases \citep{jones2008, fisher2012}. Such diseases pose a severe
threat to wildlife populations \citep{daszak2000}, and have caused
dramatic declines and extinctions in a wide range of taxa, including
echinoderms, mammals, birds, and amphibians
\citep{hewson2014, samuel2015, scheele2019, cunningham2021}. Amphibians
are experiencing particularly devastating impacts of disease due to the
recent emergence and global spread of the highly virulent amphibian
chytrid fungus, \emph{Batrachochytrium dendrobatidis} (Bd)
\citep{luedtke2023, scheele2019}. By one estimate, hundreds of species
have experienced Bd-caused declines, and numerous susceptible taxa are
extinct in the wild \citep{scheele2019}. These impacts to global
biodiversity may be unprecedented, and highlight the importance of
understanding mechanisms of species persistence in the presence of
emerging diseases \citep{russell2020}.

@@

In the current study, we had three primary objectives. First, to
determine whether the reintroduction of resistant MYL frogs obtained
from populations recovering from Bd-caused declines allows the
successful reestablishment of extirpated populations despite ongoing
disease, we conducted a 15-year landscape-scale frog reintroduction
effort (Figure~\ref{fig-recovery-model}). Second, to extend our
inferences of population recovery well beyond the temporal extent of our
reintroduction study, we developed a model to estimate the probability
of persistence for the reintroduced populations over a multi-decadal
period (Figure~\ref{fig-recovery-model}). Third, given the importance of
resistance for frog survival, population establishment, and long-term
viability (this study), we also conducted a genomic study to determine
whether MYL frogs show an evolutionary response to Bd exposure, and
whether these genomic changes are associated with resistance. To
accomplish this objective, we used exome capture methods to identify
genomic regions with consistent differences between frogs from naive
versus recovering populations (Figure~\ref{fig-recovery-model}).

\section*{Results}

\subsection*{Frog population recovery}

To determine whether MYL frogs from recovering populations can be used
to reestablish extirpated populations, we conducted 24 reintroductions
in Yosemite National Park (2006--2020). Each of the reintroductions
involved collection of adult frogs from 1 of 3 recovering, Bd-positive
``donor'' populations and translocating them to 1 of 12 nearby recipient
sites (Figure S1). The donor populations included 2 of
the 5 recovering populations included in the frog evolution study
referenced above (Figure~\ref{fig-allelefrequencies}: population 1 and
4), and these donor populations contributed frogs for 20 of the 24
translocations. Following translocation, we estimated adult survival and
recruitment of new adults from capture-mark-recapture (CMR) surveys and
obtained counts of tadpoles and juveniles from visual encounter surveys
(VES). Across all translocation sites, the duration of survey time
series was 1--16 years (median = 5).

@@

Bd loads were fairly consistent before versus after translocation, and
loads were nearly always well below the level indicative of severe
chytridiomycosis (i.e., the disease caused by Bd) and associated frog
mortality (Figure S2)
\citep{joseph2018, vredenburg2010}. Although it is possible that the
observed relatively small changes in load are a consequence of
individuals with high Bd loads dying and therefore being unavailable for
sampling during the post-translocation period, the fact that there was
little difference in pre- versus post-translocation Bd loads even in
those populations that had very high frog survival (70556, 74976;
Figure~\ref{fig-translocation-survival},
Figure S2) suggests a true lack of
substantial change in Bd load.

@@

The goal of our meta-analysis was to identify important predictors of
1-year frog survival. We were particularly interested in whether Bd load
had a negative effect on adult survival, as would be expected if frogs
were highly susceptible to Bd infection. This analysis was conducted in
a Bayesian framework and included a diversity of site, cohort, and
individual-level characteristics as predictors and 1-year frog survival
(Figure~\ref{fig-translocation-survival}) as the response variable. The
best model of 1-year frog survival identified several important
predictors, but Bd load at the time of translocation was not among them
(Figure S3). Instead, important predictors
included winter severity in the year following translocation (snow\_t1),
site elevation, and donor population (Figure~\ref{fig-cond-effects},
Figure S3). Males had somewhat higher
survival than females, but this effect was small
(Figure~\ref{fig-cond-effects}, Figure S3).
The absence of Bd load as an important predictor of frog survival is
consistent with frogs in recovering populations having sufficient
resistance to suppress Bd loads below harmful levels.

In summary, results from our frog translocation study indicate that
translocations resulted in (i) relatively high 1-year survival of
translocated adults, as well as reproduction and recruitment, at the
majority of recipient sites, (ii) 1-year survival of adults is
influenced by site characteristics, weather conditions, and donor
population (but not Bd load), and (iii) based on the relatively small
changes in Bd load after translocation, loads appear more strongly
influenced by frog characteristics (e.g., resistance) than site
characteristics. Together, these results indicate that frogs
translocated from recovering populations can maintain the benefits of
resistance in non-natal habitats. In addition, in 3 locations where
longer CMR time series allowed us to assess the survival of new adults
recruited to the population, naturally-recruited adults had equivalent
or higher survival probabilities than the originally translocated adults
(Figure S4). This suggests that frog
resistance is maintained across generations. All of the conditions
described above are supportive of population establishment and long-term
population growth.

\subsection*{Long-term population viability}

Results from the frog translocation study indicated that most
populations showed evidence of successful reproduction and recruitment,
and that adult survival was often relatively high (described above).
Although suggestive of population establishment, a decade or more of
surveys may be necessary to confirm that populations are in fact
self-sustaining \citep{joseph2018}. To extend our inferences of
population establishment beyond those possible from the site-specific
CMR data, we developed a population viability model. Specifically, to
test whether the observed yearly adult survival probabilities in
translocated populations (provided in legend of
Figure~\ref{fig-viability}B) were sufficient for long-term viability,
we built a stage-structured matrix model that captured known frog
demography and included demographic and environmental stochasticity. We
parameterized the model using CMR data from translocated populations and
known life history values in this system (Table S2
SI).

Given observed yearly adult survival probabilities of translocated frogs
(from site-specific mrmr CMR models) and a yearly survival probability
of the year-1 juvenile class (\(\sigma_{J_1}\)) greater than 0.09, at
least six of twelve translocated populations should experience a
long-run growth rate \(\lambda\) greater than 1 in the presence of Bd
(Figure~\ref{fig-viability}A; median predicted \(\lambda\) ranges from
1.19-1.40 for these six populations). These six populations all had
observed yearly adult survival greater than 0.5. As year-1 juvenile
survival probability increased above 0.2, the deterministic long-run
growth rate of eight of twelve population was greater than 1
(Figure~\ref{fig-viability}A).

Even when incorporating (i) demographic stochasticity and (ii)
environmental stochasticity in year-1 juvenile survival and recruitment
(the transition that we expect to be the most subject to environmental
variability in the presence of Bd), populations with high adult survival
are likely to persist over a 50 year time horizon. Our model predicted
that, following a single introduction of 40 adult individuals into a
population, the six populations with the highest adult survival
probabilities (\(\sigma_{A_R} > 0.5\)) had 50-year extinction
probabilities of less than 0.5 when the average year-1 juvenile survival
was greater than 0.10 (Figure~\ref{fig-viability}B). This indicates
strong potential for long-term persistence in the presence of Bd and
environmental variability in survival and recruitment. In contrast, for
the six populations where yearly adult survival probability
\(\sigma_{A_R} < 0.5\), extinction probability over 50 years was always
predicted to be \textgreater{} 50\% regardless of the value of mean
year-1 juvenile survival between 0 and 0.25. To test the validity of our
model predictions, we demonstrated that our stochastic model could
describe the general recovery trajectory of our translocated population
with the longest survey history (Figure~\ref{fig-viability}C;
population 70550, surveyed for 16 years).

In summary, our model demonstrates that given observed yearly adult
survival probabilities of translocated frogs, 50\% of our translocated
populations have a high probability of population growth and long-term
viability in the presence of Bd. This is likely a conservative estimate
because there is evidence that naturally-recruited adults have higher
survival probability than translocated adults
(Figure S4), but we considered these
probabilities to be equal in all but three of our populations where we
had sufficient data to distinguish these different probabilities.

\subsection*{Frog evolution in response to Bd}

@@

To determine whether MYL frog populations show genomic patterns
consistent with an evolutionary response to Bd, we compared frog exomes
(i.e., coding region of a genome) between populations with contrasting
histories of Bd exposure. Specifically, we compared frog genomes sampled
in 4 populations that have not yet experienced a Bd-caused epizootic
(``naive'') \citep{zhou2015} versus in 5 populations that experienced a
Bd epizootic during the past several decades and have since recovered to
varying degrees (``recovering''; Figure~\ref{fig-allelefrequencies})
\citep{knapp2016, vredenburg2010}. Bd-exposure histories of the 9 study
populations are based on 10-20 years of VES and Bd surveillance using
skin swabbing \citep[e.g.,][]{knapp2016, zhou2015, wilber2022}. Naive
populations are characterized by large numbers of adults (i.e, typically
1000s), Bd prevalence that is generally 0\% except during occasional Bd
failed invasions \citep[during which Bd loads remain very
low,][]{wilber2022}, and no history of Bd epizootics since we first
surveyed these populations in the late 1990s and early 2000s
\citep{zhou2015}. In contrast, recovering populations exist in an
enzootic state \citep{briggs2010}, characterized by smaller numbers of
adults (generally \textless{} 100), high Bd prevalence \citep[often
\textgreater{} 80\%,][]{knapp2011}, and, in adults, moderate Bd loads
that are typically well below the level expected to cause mortality
\citep{vredenburg2010}. Naive and recovering populations can be
identified unambiguously using these differences in Bd prevalence and
load. Finally, there is no potential for frog dispersal between the 9
study populations due to intervening distances and topography, as well
as the presence of introduced (predatory) fish and fish-induced habitat
fragmentation. (See \textbf{Supporting Information - Frog evolution in
response to Bd - Methods} for additional details regarding the study
design.)

@@

\hypertarget{population-viability-modeling}{%
\section*{Population viability
modeling}\label{population-viability-modeling}}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{blkarray}
\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{plainnat}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{}

\begin{document}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, enhanced, borderline west={3pt}{0pt}{shadecolor}, interior hidden, breakable, sharp corners, frame hidden]}{\end{tcolorbox}}\fi

\hypertarget{title-reintroduction-of-resistant-frogs-facilitates-landscape-scale-recovery-in-the-presence-of-a-lethal-fungal-disease}{%
\subsection{Title: Reintroduction of resistant frogs facilitates
landscape-scale recovery in the presence of a lethal fungal
disease}\label{title-reintroduction-of-resistant-frogs-facilitates-landscape-scale-recovery-in-the-presence-of-a-lethal-fungal-disease}}

\hypertarget{authors}{%
\subsubsection{Authors:}\label{authors}}

Roland A. Knapp\textsuperscript{a,b,}*, Mark Q.
Wilber\textsuperscript{c,1}, Allison Q. Byrne\textsuperscript{d,e,1},
Maxwell B. Joseph\textsuperscript{f,g}, Thomas C.
Smith\textsuperscript{a,b}, Andrew P. Rothstein\textsuperscript{d,e,h},
Robert L. Grasso\textsuperscript{i}, Erica Bree
Rosenblum\textsuperscript{d,e}

\hypertarget{affiliations}{%
\subsubsection{Affiliations:}\label{affiliations}}

\textsuperscript{a}Sierra Nevada Aquatic Research Laboratory, University
of California, Mammoth Lakes, CA, 93546

\textsuperscript{b}Earth Research Institute, University of California,
Santa Barbara, CA, 93106-3060

\textsuperscript{c}School of Natural Resources, University of Tennessee
Institute of Agriculture, Knoxville, TN, 37996

\textsuperscript{d}Department of Environmental Science, Policy, and
Management, University of California - Berkeley, Berkeley, CA,
94720-3114

\textsuperscript{e}Museum of Vertebrate Zoology, University of
California - Berkeley, Berkeley, CA, 94720-3160

\textsuperscript{f}Earth Lab, University of Colorado, Boulder, CO, 80303

\textsuperscript{g}Current affiliation: Planet, San Francisco, CA, 94107

\textsuperscript{h}Current affiliation: Ginkgo Bioworks, Boston, MA,
02210

\textsuperscript{i}Resources Management and Science, Yosemite National
Park, El Portal, CA, 95318

*Corresponding author: Sierra Nevada Aquatic Research Laboratory, 1016
Mount Morrison Road, Mammoth Lakes, CA 93546, phone: 760.647.0034,
email: roland.knapp@ucsb.edu

\textsuperscript{1}These authors contributed equally to this work.

\textbf{Classification:} Biological Sciences - Ecology

\textbf{Key words:} evolution, resistance, amphibians, population
recovery, reintroduction, disease, \emph{Batrachochytrium dendrobatidis}

\hypertarget{abstract}{%
\subsubsection{Abstract}\label{abstract}}

Vast alteration of the biosphere by humans is causing a sixth mass
extinction, driven in part by an increase in emerging infectious
diseases. The emergence of the lethal fungal pathogen
(\emph{Batrachochytrium dendrobatidis}; ``Bd'') has devastated global
amphibian biodiversity, with hundreds of species experiencing declines
or extinctions. With no broadly applicable methods available to reverse
these impacts in the wild, the future of many amphibians appears grim.
The once-common mountain yellow-legged (MYL) frog is emblematic of
amphibians threatened by Bd. Although most MYL frog populations are
extirpated following disease outbreaks, some persist and eventually
recover. Frogs in these recovering populations have increased resistance
against Bd infection, consistent with evolution of resistant genotypes
and/or acquired immunity. We conducted a 15-year landscape-scale
reintroduction study and show that frogs collected from recovering
populations and reintroduced to vacant habitats can reestablish
populations despite the presence of Bd. In addition, results from
viability modeling suggest that many reintroduced populations have a low
probability of extinction over 50 years. To better understand the role
of evolution in frog resistance, we compared the genomes of MYL frogs
from Bd-naive and recovering populations. We found substantial
differences between these categories, including changes in immune
function loci that may confer increased resistance, consistent with
evolutionary changes in response to Bd exposure. These results provide a
rare example of how reintroduction of resistant individuals can allow
the landscape-scale recovery of disease-impacted species. This example
has broad implications for the many taxa worldwide that are threatened
with extinction by novel pathogens.

\hypertarget{significance}{%
\subsubsection{Significance}\label{significance}}

Understanding how species persist despite accelerating global change is
critical for the conservation of biodiversity. Emerging infectious
diseases can have particularly devastating impacts, and few options
exist to reverse these effects. We used large-scale reintroductions of
disease-resistant individuals in an effort to recover a once-common frog
species driven to near-extinction by a disease that has decimated
amphibian biodiversity. Introduction of resistant frogs allowed
reestablishment of viable populations in the presence of disease. In
addition, resistance may be at least partially the result of natural
selection at specific immune function genes, which show evidence for
selection in recovering populations. The evolution of resistance and
reintroduction of resistant individuals could play an important role in
biodiversity conservation in our rapidly changing world.

\hypertarget{introduction}{%
\subsubsection{Introduction}\label{introduction}}

Human activities are increasingly impacting global biodiversity
\citep{ceballos2015}, with important implications for ecosystem
resilience and human welfare \citep{naeem2009}. One consequence of human
alteration of the biosphere is an increase in emerging infectious
diseases \citep{jones2008, fisher2012}. Such diseases pose a severe
threat to wildlife populations \citep{daszak2000}, and have caused
dramatic declines and extinctions in a wide range of taxa, including
echinoderms, mammals, birds, and amphibians
\citep{hewson2014, samuel2015, scheele2019, cunningham2021}. Amphibians
are experiencing particularly devastating impacts of disease due to the
recent emergence and global spread of the highly virulent amphibian
chytrid fungus, \emph{Batrachochytrium dendrobatidis} (Bd)
\citep{luedtke2023, scheele2019}. By one estimate, hundreds of species
have experienced Bd-caused declines, and numerous susceptible taxa are
extinct in the wild \citep{scheele2019}. These impacts to global
biodiversity may be unprecedented, and highlight the importance of
understanding mechanisms of species persistence in the presence of
emerging diseases \citep{russell2020}.

@@

In the current study, we had three primary objectives. First, to
determine whether the reintroduction of resistant MYL frogs obtained
from populations recovering from Bd-caused declines allows the
successful reestablishment of extirpated populations despite ongoing
disease, we conducted a 15-year landscape-scale frog reintroduction
effort (Figure~\ref{fig-recovery-model}). Second, to extend our
inferences of population recovery well beyond the temporal extent of our
reintroduction study, we developed a model to estimate the probability
of persistence for the reintroduced populations over a multi-decadal
period (Figure~\ref{fig-recovery-model}). Third, given the importance of
resistance for frog survival, population establishment, and long-term
viability (this study), we conducted a genomic study using exome capture
methods to determine whether MYL frogs in recovering populations show
evidence of selection and whether these genomic changes are associated
with resistance (Figure~\ref{fig-recovery-model}).

\hypertarget{results}{%
\subsubsection{Results}\label{results}}

\hypertarget{frog-population-recovery}{%
\paragraph{Frog population recovery}\label{frog-population-recovery}}

To determine whether MYL frogs from recovering populations can be used
to reestablish extirpated populations, we conducted 24 reintroductions
in Yosemite National Park (2006--2020). Each of the reintroductions
involved collection of adult frogs from 1 of 3 recovering, Bd-positive
``donor'' populations and translocating them to 1 of 12 nearby recipient
sites (Figure~\ref{fig-yosemap} SI). The donor populations included 2 of
the 5 recovering populations used in the frog evolution study referenced
above (Figure~\ref{fig-allelefrequencies}: population 1 and 4), and
these donor populations contributed frogs for 20 of the 24
translocations. Following translocation, we estimated adult survival and
recruitment of new adults from capture-mark-recapture (CMR) surveys and
obtained counts of tadpoles and juveniles from visual encounter surveys
(VES). Across all translocation sites, the duration of survey time
series was 1--16 years (median = 5).

@@

Bd loads were fairly consistent before versus after translocation, and
loads were nearly always well below the level indicative of severe
chytridiomycosis (i.e., the disease caused by Bd) and associated frog
mortality (Figure~\ref{fig-bdload-beforeafter} SI)
\citep{joseph2018, vredenburg2010}. Although it is possible that the
observed relatively small changes in load are a consequence of
individuals with high Bd loads dying and therefore being unavailable for
sampling during the post-translocation period, the fact that there was
little difference in pre- versus post-translocation Bd loads even in
those populations that had very high frog survival (70556, 74976 - see
below; Figure~\ref{fig-bdload-beforeafter} SI) suggests a true lack of
substantial change in Bd load.

@@

The goal of our meta-analysis was to identify important predictors of
1-year frog survival. We were particularly interested in whether Bd load
had a negative effect on adult survival, as would be expected if frogs
were highly susceptible to Bd infection. This analysis was conducted in
a Bayesian framework and included a diversity of site, cohort, and
individual-level characteristics as predictors and 1-year frog survival
(Figure~\ref{fig-translocation-survival}) as the response variable. The
best model of 1-year frog survival identified several important
predictors, but Bd load at the time of translocation was not among them
(Figure~\ref{fig-survival-postdens} SI). Instead, important predictors
included winter severity in the year following translocation (snow\_t1),
site elevation, and donor population (Figure~\ref{fig-cond-effects},
Figure~\ref{fig-survival-postdens} SI). Males had somewhat higher
survival than females, but this effect was small
(Figure~\ref{fig-cond-effects}, Figure~\ref{fig-survival-postdens} SI).
The absence of Bd load as an important predictor of frog survival is
consistent with frogs in recovering populations having sufficient
resistance to suppress Bd loads below harmful levels.

In summary, results from our frog translocation study indicate that
translocations resulted in (i) relatively high 1-year survival of
translocated adults, as well as reproduction and recruitment, at the
majority of recipient sites, (ii) 1-year survival of adults is
influenced by site characteristics, weather conditions, and donor
population (but not Bd load), and (iii) based on the relatively small
changes in Bd load after translocation, loads appear more strongly
influenced by frog characteristics (e.g., resistance) than site
characteristics. Together, these results indicate that frogs
translocated from recovering populations can maintain the benefits of
resistance in non-natal habitats. In addition, in 3 locations where
longer CMR time series allowed us to assess the survival of new adults
recruited to the population, naturally-recruited adults had equivalent
or higher survival probabilities than the originally translocated adults
(Figure~\ref{fig-compare_surv_probs} SI). This suggests that frog
resistance is maintained across generations. All of the conditions
described above are supportive of population establishment and long-term
population growth.

\hypertarget{long-term-population-viability}{%
\paragraph{Long-term population
viability}\label{long-term-population-viability}}

Results from the frog translocation study indicated that most
populations showed evidence of successful reproduction and recruitment,
and that adult survival was often relatively high (described above).
Although suggestive of population establishment, a decade or more of
surveys may be necessary to confirm that populations are in fact
self-sustaining \citep{joseph2018}. To extend our inferences of
population establishment beyond those possible from the site-specific
CMR data, we developed a population viability model. Specifically, to
test whether the observed yearly adult survival probabilities in
translocated populations were sufficient for long-term viability, we
built a stage-structured matrix model that captured known frog
demography and included demographic and environmental stochasticity. We
parameterized the model using CMR data from translocated populations and
known life history values in this system (Table~\ref{tbl-param_values}
SI).

Given observed yearly adult survival probabilities of translocated frogs
(from site-specific mrmr CMR models; provided in legend of
Figure~\ref{fig-viability} B) and a yearly survival probability of the
year-1 juvenile class (\(\sigma_{J_1}\)) greater than 0.09, at least six
of twelve translocated populations should experience a long-run growth
rate \(\lambda\) greater than 1 in the presence of Bd
(Figure~\ref{fig-viability} A; median predicted \(\lambda\) ranges from
1.19-1.40 for these six populations). These six populations all had
observed yearly adult survival greater than 0.5. As year-1 juvenile
survival probability increased above 0.2, the deterministic long-run
growth rate of eight of twelve population was greater than 1
(Figure~\ref{fig-viability} A).

Even when incorporating (i) demographic stochasticity and (ii)
environmental stochasticity in year-1 juvenile survival and recruitment
(the transition that we expect to be the most subject to environmental
variability in the presence of Bd), populations with high adult survival
are likely to persist over a 50 year time horizon. Our model predicted
that, following a single introduction of 40 adult individuals into a
population, the six populations with the highest adult survival
probabilities (\(\sigma_{A_R} > 0.5\)) had 50-year extinction
probabilities of less than 0.5 when the average year-1 juvenile survival
was greater than 0.10 (Figure~\ref{fig-viability} B). This indicates
strong potential for long-term persistence in the presence of Bd and
environmental variability in survival and recruitment. In contrast, for
the six populations where yearly adult survival probability
\(\sigma_{A_R} < 0.5\), extinction probability over 50 years was always
predicted to be \textgreater{} 50\% regardless of the value of mean
year-1 juvenile survival between 0 and 0.25. To test the validity of our
model predictions, we demonstrated that our stochastic model could
describe the general recovery trajectory of our translocated population
with the longest survey history (Figure~\ref{fig-viability} C;
population 70550, surveyed for 16 years).

In summary, our model demonstrates that given observed yearly adult
survival probabilities of translocated frogs, 50\% of our translocated
populations have a high probability of population growth and long-term
viability in the presence of Bd. This is likely a conservative estimate
because there is evidence that naturally-recruited adults have higher
survival probability than translocated adults
(Figure~\ref{fig-compare_surv_probs} SI), but we considered these
probabilities to be equal in all but three of our populations where we
had sufficient data to distinguish these different probabilities.

\hypertarget{frog-evolution-in-response-to-bd}{%
\paragraph{Frog evolution in response to
Bd}\label{frog-evolution-in-response-to-bd}}

@@

To determine whether MYL frog populations show genomic patterns
consistent with an evolutionary response to Bd, we compared frog exomes
(i.e., coding region of a genome) between populations with contrasting
histories of Bd exposure. Specifically, we compared frog genomes sampled
in 4 populations that have not yet experienced a Bd-caused epizootic
(``naive'') \citep{zhou2015} versus in 5 populations that experienced a
Bd epizootic during the past several decades and have since recovered to
varying degrees (``recovering''; Figure~\ref{fig-allelefrequencies})
\citep{knapp2016, vredenburg2010}. Bd-exposure histories of the 9 study
populations are based on 10-20 years of VES and Bd surveillance using
skin swabbing \citep[e.g.,][]{knapp2016, zhou2015, wilber2022}. Naive
populations are characterized by large numbers of adults (i.e, typically
1000s), Bd prevalence that is generally 0\% except during occasional Bd
failed invasions \citep[during which Bd loads remain very
low,][]{wilber2022}, and no history of Bd epizootics since we first
surveyed these populations in the late 1990s and early 2000s
\citep{zhou2015}. In contrast, recovering populations exist in an
enzootic state \citep{briggs2010}, characterized by smaller numbers of
adults (generally \textless{} 500), high Bd prevalence \citep[often
\textgreater{} 80\%,][]{knapp2011}, and, in adults, moderate Bd loads
that are typically well below the level expected to cause mortality
\citep{vredenburg2010}. Naive and recovering populations can be
identified unambiguously using these differences in Bd prevalence and
load. Finally, there is no potential for frog dispersal between the 9
study populations due to intervening distances and topography, as well
as the presence of introduced (predatory) fish and fish-induced habitat
fragmentation. (See \textbf{Supporting Information - Frog evolution in
response to Bd - Study design} for additional details regarding the
study design.)

@@

\hypertarget{population-viability-modeling}{%
\paragraph{Population viability
modeling}\label{population-viability-modeling}}
